Hydrodynamics of a Semipermeable Inextensible Membrane Under Flow and Confinement

### 2021-05-18 vesicle_code/docs/permeableVes/Response2Refs.tex
worked on revisions

\newpage
\vspace{0.5cm}
{\large \bf Response to Referee 1:}
\vspace{0.5cm}

@@

\noindent
Response: We thank the referee for this point. In our formulation, the permeation flux is proportional to the hydrostatic pressure and the normal component of the membrane force, which is the capillary pressure due to the Lagrange multiplier tension. The tangential force component due to tension field does not contribute to the permeation flux. To elucidate the relative contribution between the hydrostatic pressure and the normal component of the membrane force, we have added a short paragraph to illustrate how the relative contributions are in all cases.

@@

\newpage
\vspace{0.5cm}
{\large \bf Response to Referee 2:}
\vspace{0.5cm}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\vspace{0.5cm}
{\large \bf Response to Referee 1:}
\vspace{0.5cm}

@@

\noindent
Response: We thank the referee for this point. In our formulation, the
    permeation flux is proportional to the normal component of the
    membrane force, which is also the capillary pressure due to both
    bending and the Lagrange multiplier tension. The tangential
    force component due to tension field does not contribute to the
    permeation flux. To elucidate the relative contribution between the
    hydrostatic pressure due to the bending and tension forces, we have
    added additional figures and short paragraphs to illustrate the
    relative contributions are in all cases.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\vspace{0.5cm}
{\large \bf Response to Referee 2:}
\vspace{0.5cm}

### 2021-05-18 vesicle_code/docs/permeableVes/semipermeable_vesicle.tex
worked on revisions

We next compute the areal dynamics of a semipermeable vesicle. Since the
area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term is zero because of the local inextensibility condition.
Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals are identical and we combine them to obtain
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Substituting in $\ff = \ff_\mathrm{ben} + \ff_\mathrm{ten}$ and
repeatedly applying integration by parts, we obtain
\begin{align}
  \dot{A}(t) & = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
%\begin{align}
%  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
%    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
%  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
%    \right) \cdot \kappa \xx_s \, ds \\
%  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
%    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
%  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
%    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
%    \, ds \\
%  &  = -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
%    \, ds = \beta \int_\gamma \left(
%    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
%  \label{eqn:areaROC}
%\end{align}
Note that the integrand in Equation~\eqref{eqn:areaROC} is zero when the membrane tension balances the elastic force
\begin{align}
\label{eqn:areaROC2}
\Lambda &= \frac{\kappa^2}{2}=\frac{\kappa_0^2}{2},
%  \;\mathrm{ at }\; \mathrm{equilibrium},
\end{align}
which is consistent with the equilibrium result in
Equation~(\ref{eqn:curveODE}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hydrodynamics of a semipermeable inextensible membrane in free space\label{sec:free_space}}
To highlight the effects of membrane permeability to water, we consider
a single semipermeable inextensible membrane, with (dimensionless) permeability
coefficient $\beta>0$, in familiar configurations: In free space we
place a semipermeable inextensible membrane in a quiescent flow, a planar shear flow,
and a Poiseuille flow. For each example, we fix the membrane length and
examine the enclosed water content as a dynamic consequence of water
flow, membrane shape deformation, and distribution of membrane tension.
%
Under these flows, our simulations show that the steady state of an
impermeable vesicle is different from that of a semipermeable inextensible membrane,
which depends only on the strength of the external flow and is
insensitive to permeability $\beta$ (as long as $\beta>0$). We find that
the time in the dynamic evolution towards equilibrium is scaled by
$\beta^{-1}$: the smaller $\beta$ is the longer it takes to reach the
steady state.

@@

@@

\section{Hydrodynamics of A semipermeable inextensible membrane under strong confinement\label{sec:confinement}}
Here we consider a semipermeable inextensible membrane under strong
confinement: a long closely-fit channel and a contracting channel. In a
narrow microfluidic channel, a large pressure jump is often used to push
the vesicle through a closely-fit channel~\cite{abk-fai-sto2006}. Our
simulations show that such a configuration can lead to an amplification
of water permeation at a fast time scale of the order of milliseconds.
In addition, we find that a large value of $\beta$ is needed to keep the
membrane tension below the threshold value for membrane poration. Such
high permeability for water may be regarded as an indication that the
lipid bilayer membrane is porated under strong confinement, and our
inextensible membrane model is for a porated vesicle within the Helfrich
free energy framework. As in the previous examples, we fix the membrane
length and examine the enclosed water content as a dynamic consequence
of water flow, membrane shape deformation, and membrane tension
distribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Semipermeable Inextensible Membrane in a Closely-fit Channel (Stenosis)}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/stenosisComposite.tikz}
  \else
  \includegraphics{figures/stenosisComposite.pdf}
  \fi
  \caption{\label{fig:stenosisComposite} (a) A semipermeable
  inextensible membrane passing through a closely-fit channel similar to
  the experimental device in~\citet{abk-fai-sto2006}. (b) A
  semipermeable inextensible membrane passing through a (three-times)
  longer closely-fit channel than (a). In (a) and (b), red regions
  correspond to influx and blue regions correspond to efflux. (c) and
  (d) The tension in N/m of the vesicle shapes in (a) and (b). (e) The
  excess pressure of an impermeable and semipermeable inextensible
  membrane in the short geometry in (a). (f) The reduced area in the
  geometries in (a) and (b) as a function of the membranes' center of
  mass. In (e) and (f), the x-axis is scaled by $1/L$ where $L$ is half
  the length of the constriction. (g) The reduced area in the geometries
  in (a) and (b) as a function of time. The time axis is scaled by
  $1/t_R$ where $t_R$ is the residency time. In (e) and (f), the dashed
  lines correspond to the locations of the inlet and outlet. In (g), the
  dashed lines correspond to the times that the vesicle enters and exits
  the constriction.}
\end{figure*}
When red blood cells go through small capillary vessels, they experience
large shape deformation and significant membrane stretching that
triggers ATP release~\cite{Wan2008_PNAS, ForsythWan2011_PNAS}. Many
cells exhibit extraordinary flexibility as they go through narrow
vessels~\cite{AuStoreyMoore2016_PNAS}, and adjustment of
surface-to-volume ratio may be a key factor for a successful passage.
Here we quantify the effect of water permeability on a vesicle going
through a stenosis (a long microfluidic channel with a width smaller
than the cell radius). Based on previous
findings~\cite{abk-fai-sto2006}, we first examine the effect of
semipermeability on the pressure fluctuation across a stenosed geometry.
In this experiment, a pressure drop is used to squeeze a red blood cell
through a constriction with diameter $5\;\mu$m and length $45\;\mu$m. We
set up a similar computational geometry with a Poiseuille flow imposed
at the inlet and outlet (Figure~\ref{fig:stenosisComposite}(a)). We set
the maximum flow velocity to $U_{\max} = 1000 \;\mu$m/s at the inlet and
outlet so that the residency time of the membrane in the stenosis
channel is $t_R \approx 25$ ms which agrees with the experimental
results~\cite{abk-fai-sto2006}. When the red blood cell enters the
channel, an increase in the pressure drop is required to maintain a
constant flow rate, and the difference in the pressure drop is called
the excess pressure.

@@

%When a semipermeable vesicle goes through a closely-fit channel, we
%observe that the water efflux is found in the front, and the influx is
%in the rear of the vesicle membrane
%(Figure~\ref{fig:stenosisComposite}). This corresponds well to the
%corresponding tension distribution in Figure~\ref{fig:stenosisTensions}:
%Larger tension in the front, and smaller tension in the rear.

@@

>>>>

We next compute the areal dynamics of a semipermeable vesicle. Since the
area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term is zero because of the local inextensibility condition.
Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals are identical and we combine them to obtain
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
  \label{eqn:areaROCforce}
\end{align}
where we have used the incompressibility of the single-layer potential.
Substituting in $\ff = \ff_\mathrm{ben} + \ff_\mathrm{ten}$ and
repeatedly applying integration by parts, we obtain
\begin{align}
  \dot{A}(t) & = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
%\begin{align}
%  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
%    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
%  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
%    \right) \cdot \kappa \xx_s \, ds \\
%  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
%    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
%  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
%    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
%    \, ds \\
%  &  = -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
%    \, ds = \beta \int_\gamma \left(
%    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
%  \label{eqn:areaROC}
%\end{align}
Note that the integrand in Equation~\eqref{eqn:areaROC} is zero when the membrane tension balances the elastic force
\begin{align}
\label{eqn:areaROC2}
\Lambda &= \frac{\kappa^2}{2}=\frac{\kappa_0^2}{2},
%  \;\mathrm{ at }\; \mathrm{equilibrium},
\end{align}
which is consistent with the equilibrium result in
Equation~(\ref{eqn:curveODE}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hydrodynamics of a Semipermeable Inextensible Membrane in Free Space\label{sec:free_space}}
To highlight the effects of membrane permeability to water, we consider
a single semipermeable inextensible membrane, with (dimensionless) permeability
coefficient $\beta>0$, in familiar configurations: In free space we
place a semipermeable inextensible membrane in a quiescent flow, a planar shear flow,
and a Poiseuille flow. For each example, we fix the membrane length and
examine the enclosed water content as a dynamic consequence of water
flow, membrane shape deformation, and distribution of membrane tension.
%
Under these flows, our simulations show that the steady state of an
impermeable vesicle is different from that of a semipermeable inextensible membrane,
which depends only on the strength of the external flow and is
insensitive to permeability $\beta$ (as long as $\beta>0$). We find that
the time in the dynamic evolution towards equilibrium is scaled by
$\beta^{-1}$: the smaller $\beta$ is the longer it takes to reach the
steady state.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Role of Tension and Bending on Semipermeability}
\textcolor{magenta}{Since $\jump{p}= \ff_\mathrm{mem} \cdot
\nn$~\cite{poz1992}, our model assumes that the permeating flux is
proportional to the mechanical pressure jump across the membrane. We
decompose the normal component of the membrane force into its
contributions due to tension and bending as
\begin{align}
  \ff_{\mathrm{ten}} \cdot \nn = -\kappa \Lambda 
    \quad \text{ and } \quad
  \ff_{\mathrm{ben}} \cdot \nn = \kappa_{ss} + \frac{1}{2}\kappa^3.
\end{align}
Then, the total flux through the membrane due to the tension and bending
forces are
\begin{align}
  \dot{A}_{\mathrm{ten}}(t) = -\beta \int_{\gamma} \kappa \Lambda\, ds  
  \quad \text{ and } \quad
  \dot{A}_{\mathrm{ben}}(t) = \frac{\beta}{2} \int_{\gamma} \kappa^3\, ds,
  \label{eqn:areaROCdecomposed}
\end{align}
respectively. Note that the sum of $\dot{A}_{\mathrm{ten}}(t)$ and
$\dot{A}_{\mathrm{ben}}(t)$ results in equation~\eqref{eqn:areaROC}.
After multiplying~\eqref{eqn:areaROCdecomposed} by $4\pi/L^2$, where $L$
is the constant  vesicle length, the rate of change of the reduced area
due to the tension and bending forces are formed. We investigate the
effect of the tension and bending forces on the semipermeability for two
quiescent examples (Figures~\ref{fig:unconfinedRAdecomp}(a) and (b)),
one planar shear example (Figure~\ref{fig:unconfinedRAdecomp}(c)), and
one Poiseuille example (Figure~\ref{fig:unconfinedRAdecomp}(d)). In the
quiescent flow examples, we see that the tension and bending forces
result in similar amounts of semipermeability, and for the geometry with
a larger initial curvature, the reduced area changes much quicker and
the vesicle deflates at early times before inflating to a circular
shape. In the planar shear flow and Poiseuille flow examples, the
tension and bending forces result in similar amounts of
semipermeability.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/unconfinedRAdecomp.tikz}
  \else
  \includegraphics{figures/unconfinedRAdecomp.pdf}
  \fi
  \caption{\label{fig:unconfinedRAdecomp}\textcolor{magenta}{The rate of
  change of the reduced area caused by the tension and bending forces.
  The dashed line is the sum of these two contributions. (a) The
  quiescent example in Figure~\ref{fig:relaxationComposite} with
  $\alpha(0) = 0.66$. (b) The quiescent example in
  Figure~\ref{fig:relaxationComposite} with $\alpha(0) = 0.38$. (c) The
  planar shear example in Figure~\ref{fig:shearComposite} with $\alpha(0) =
  0.45$ and $Ca_\mathrm{E} = 10^0$. (d) The Poiseuille example in
  Figure~\ref{fig:parabolicComposite} with $\alpha(0) = 0.60$ and
  maximum velocity $U = 600~\mu$m/s.}}
\end{figure*}

@@

\section{Hydrodynamics of a Semipermeable Inextensible Membrane Under Strong Confinement\label{sec:confinement}}
Here we consider a semipermeable inextensible membrane under strong
confinement: a long closely-fit channel and a contracting channel. In a
narrow microfluidic channel, a large pressure jump is often used to push
the vesicle through a closely-fit channel~\cite{abk-fai-sto2006}. Our
simulations show that such a configuration can lead to an amplification
of water permeation at a fast time scale of the order of milliseconds.
In addition, we find that a large value of $\beta$ is needed to keep the
membrane tension below the threshold value for membrane poration. Such
high permeability for water may be regarded as an indication that the
lipid bilayer membrane is porated under strong confinement, and our
inextensible membrane model is for a porated vesicle within the Helfrich
free energy framework. As in the previous examples, we fix the membrane
length and examine the enclosed water content as a dynamic consequence
of water flow, membrane shape deformation, and membrane tension
distribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Semipermeable Inextensible Membrane in a Closely-fit
Channel (Stenosis)}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/stenosisComposite.tikz}
  \else
  \includegraphics{figures/stenosisComposite.pdf}
  \fi
  \caption{\label{fig:stenosisComposite} (a) A semipermeable
  inextensible membrane passing through a closely-fit channel similar to
  the experimental device in~\citet{abk-fai-sto2006}. (b) A
  semipermeable inextensible membrane passing through a (three-times)
  longer closely-fit channel than (a). In (a) and (b), red regions
  correspond to influx and blue regions correspond to efflux. (c) and
  (d) The tension in N/m of the vesicle shapes in (a) and (b). (e) The
  excess pressure of an impermeable and semipermeable inextensible
  membrane in the short geometry in (a). (f) The reduced area in the
  geometries in (a) and (b) as a function of the membranes' center of
  mass. In (e) and (f), the x-axis is scaled by $1/L$ where $L$ is half
  the length of the constriction. (g) The reduced area in the geometries
  in (a) and (b) as a function of time. The time axis is scaled by
  $1/t_R$ where $t_R$ is the residency time. In (e) and (f), the dashed
  lines correspond to the locations of the inlet and outlet. In (g), the
  dashed lines correspond to the times that the vesicle enters and exits
  the constriction.}
\end{figure*}

When red blood cells go through small capillary vessels, they experience
large shape deformation and significant membrane stretching that
triggers ATP release~\cite{Wan2008_PNAS, ForsythWan2011_PNAS}. Many
cells exhibit extraordinary flexibility as they go through narrow
vessels~\cite{AuStoreyMoore2016_PNAS}, and adjustment of
surface-to-volume ratio may be a key factor for a successful passage.
Here we quantify the effect of water permeability on a vesicle going
through a stenosis (a long microfluidic channel with a width smaller
than the cell radius). Based on previous
findings~\cite{abk-fai-sto2006}, we first examine the effect of
semipermeability on the pressure fluctuation across a stenosed geometry.
In this experiment, a pressure drop is used to squeeze a red blood cell
through a constriction with diameter $5\;\mu$m and length $45\;\mu$m. We
set up a similar computational geometry with a Poiseuille flow imposed
at the inlet and outlet (Figure~\ref{fig:stenosisComposite}(a)). We set
the maximum flow velocity to $U_{\max} = 1000 \;\mu$m/s at the inlet and
outlet so that the residency time of the membrane in the stenosis
channel is $t_R \approx 25$ ms which agrees with the experimental
results~\cite{abk-fai-sto2006}. When the red blood cell enters the
channel, an increase in the pressure drop is required to maintain a
constant flow rate, and the difference in the pressure drop is called
the excess pressure.

@@

\textcolor{magenta}{Using equation~\eqref{eqn:areaROCdecomposed}, we
investigate the effects of the tension and bending forces on the water
flux through the membrane. In Figure~\ref{fig:stenosisRAdecomp}, we
compute the rate of change of the reduced area due to the tension force
(blue), the bending force (red), and the total membrane force (black).
In both the short and long closely-fit channels, the tension force is
responsible for nearly all the semipermeability. This is in contrast to
the unconfined examples where the tension and bending forces were
each responsible for a comparable amount of semipermeability.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/stenosisRAdecomp.tikz}
  \else
  \includegraphics{figures/stenosisRAdecomp.pdf}
  \fi
  \caption{\label{fig:stenosisRAdecomp} \textcolor{magenta}{The rate of
  change of the reduced area caused by the tension and bending forces.
  The dashed line is the sum of these two contributions. (a) The short
  stenosis geometry in Figure~\ref{fig:stenosisComposite}(a). (b) The
  long stenosis geometry in Figure~\ref{fig:stenosisComposite}(b).}}
\end{figure*}

@@

\textcolor{magenta}{As was done for the closely-fit channel, we use
equation~\eqref{eqn:areaROCdecomposed} to investigate the effects of the
tension and bending forces on semipermeability.
Figure~\ref{fig:contractingRAdecomp} shows the rate of change of the
reduced area due to the tension force (blue), the bending force (red),
and the total membrane force (black). Again, the tension is responsible
for the majority of the semipermeability. However, when compared to the
closely-fit channel example, the bending does have more of a
contribution because of the lower maximum imposed velocity. While we
only decompose the membrane force into tension and bending for the two
vesicles that successfully pass through the contraction, we also plot
the total membrane force for the two vesicles that do not pass through
the contraction (green). The vesicle does deflate with a smaller
semipermeability coefficient and a smaller maximum imposed velocity, but
the rate of efflux is insufficient for successful passage of the vesicle
through the contraction.}

\begin{figure*}[htp]
  \ifTikz
  \input{figures/contractingRAdecomp.tikz}
  \else
  \includegraphics{figures/contractingRAdecomp.pdf}
  \fi
  \caption{\label{fig:contractingRAdecomp} \textcolor{magenta}{The rate
  of change of the reduced area caused by the tension and bending
  forces. The dashed line is the sum of these two contributions. Only
  simulations where the vesicle passes through the contraction are
  decomposed into the contributions due to tension and bending. (a) The
  low semipermeability coefficient in
  Figure~\ref{fig:contractingComposite}(b). The green line corresponds
  to the simulation in Figure~\ref{fig:contractingComposite}(a). (b) The
  high semipermeability coefficient in
  Figure~\ref{fig:contractingComposite}(e). The green line corresponds
  to the simulation in Figure~\ref{fig:contractingComposite}(d).}}
\end{figure*}
